Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-06 MDPI_submission.tex
Commit 2023-12-06 23:10

We calculated the linear SDR trends, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are very similar and are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
smaller trend for clear-skies (\(0.024\,\%/y\)) and a negative weaker
trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR
slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
before the start of our records. It is interesting to note, that for the
observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
thus we can not attribute any major effect on SDR trend to Solar
variability.

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend p-value & Pearson cor. & Days\\
\midrule
All skies & 0.3756 & 0.000 & 0.0911 & 10250\\
Clear skies & 0.0239 & 0.189 & 0.0172 & 5803\\
Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
\bottomrule
\end{tabu}
\end{table}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
\midrule
\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
\cmidrule{1-7}
0.1360 & 2.19e-01 & 0.0515 & 0.2270307 & All skies & Spring & 78.1\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
\cmidrule{1-7}
0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
-0.1450 & 1.49e-01 & 0.0706 & -0.2656359 & Clear skies & Spring & 85.1\\
\cmidrule{1-7}
\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
\cmidrule{1-7}
0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
\cmidrule{1-7}
0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
\bottomrule
\end{tabu}
\end{table}

>>>>

We calculated the linear SDR trends, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are very similar and are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
smaller trend for clear-skies (\(0.024\,\%/y\)) and a negative weaker
trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR
slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
before the start of our records. It is interesting to note, that for the
observations period, the trend of the TSI is \(-0.00022\,\%/y\), and
thus we can not attribute any major effect on SDR trend to Solar
variability.

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend p-value & Pearson cor. & Days\\
\midrule
All skies & 0.3765 & 0.000 & 0.0913 & 10251\\
Clear skies & 0.0244 & 0.179 & 0.0176 & 5804\\
Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
\bottomrule
\end{tabu}
\end{table}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & cor.estimate & cor.conf\_level & Sky condition & Season & slope.stat\_sig\\
\midrule
\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
\cmidrule{1-8}
0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & 0.95 & All skies & Spring & 82.2\\
\cmidrule{1-8}
\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
\cmidrule{1-8}
0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & 0.95 & All skies & Autumn & 99.3\\
\cmidrule{1-8}
\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-8}
-0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & 0.95 & Clear skies & Spring & 82.9\\
\cmidrule{1-8}
\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-8}
0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & 0.95 & Clear skies & Autumn & 42.1\\
\cmidrule{1-8}
\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
\cmidrule{1-8}
0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & 0.95 & Cloudy skies & Spring & 87.5\\
\cmidrule{1-8}
\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
\cmidrule{1-8}
0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & 0.95 & Cloudy skies & Autumn & 97.5\\
\bottomrule
\end{tabu}
\end{table}
